---
title: "exampletestr"
output: github_document
---

```{r knitr setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

An R package to help developers create unit tests (designed for use with the testthat package) for their package, based on the examples in their package documentation. The documentation must be done with roxygen.

[![Travis-CI Build Status](https://travis-ci.org/rorynolan/exampletestr.svg?branch=master)](https://travis-ci.org/rorynolan/exampletestr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/rorynolan/exampletestr?branch=master&svg=true)](https://ci.appveyor.com/project/rorynolan/exampletestr)
[![codecov](https://codecov.io/gh/rorynolan/exampletestr/branch/master/graph/badge.svg)](https://codecov.io/gh/rorynolan/exampletestr)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/exampletestr)](https://cran.r-project.org/package=exampletestr)
![RStudio CRAN downloads](http://cranlogs.r-pkg.org/badges/grand-total/exampletestr)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

## Installation
In R, enter
```{r Install exampletestr, eval=FALSE}
install.packages("devtools")
devtools::install_github("rorynolan/exampletestr")
```
and you're done!

## Use
First, let's set up a dummy package directory with just the `utils.R` file from the source code of the `exampletestr` package.
```{r setup}
library(exampletestr)
library(testthat)
devtools::create("tempkg")
devtools::use_testthat("tempkg")
file.copy(system.file("extdata", "utils.R", package = "exampletestr"), 
          "tempkg/R")
devtools::document("tempkg")
```

The `utils.R` file looks like this:
```{r Look at utils file, eval=FALSE}
#' Does parsing of text give an error?
#'
#' Can a character vector (where each line is treated as a line of R code) be
#' parsed as an R expression (or several R expressions) without giving an
#' error?
#'
#' @param text_expr The expression to be evaluated, as a character vector.
#'
#' @return `TRUE` if the code gives an error and `FALSE` otherwise.
#' @examples
#' text_parse_error("a <- 1")
#' text_parse_error("a <- ")
#' @export
text_parse_error <- function(text_expr) {
  try_res <- try(parse(text = text_expr), silent = TRUE)
  error <- inherits(try_res, "try-error")
  if (error) attr(error, "message") <- attr(try_res, "message")
  error
}

#' Text expression groups.
#'
#' Given a character vector of R expressions, break the vector up into groups of
#' lines, where each group of lines is a valid R expression.
#'
#' @param text_expr A character vector.
#' @param remove_comments Should comments be removed?
#'
#' @return A list of character vectors, each of which can be evaluated as a
#'   valid R expression.
#' @examples
#' text_expr <- c("a <- 1",
#' "fx <- function(x) {",
#' "  x + 1",
#' "}  # this comment should disappear")
#' extract_expressions(text_expr)
#' @export
extract_expressions <- function(text_expr, remove_comments = TRUE) {
  stopifnot(length(text_expr) > 0)
  expr_groups <- list()
  i <- 1
  while (i <= length(text_expr)) {
    j <- 0
    expr <- text_expr[i]
    while(text_parse_error(expr)) {
      j <- j + 1
      expr <- text_expr[i:(i + j)]
    }
    expr_groups <- append(expr_groups, list(expr))
    i <- i + j + 1
  }
  if (remove_comments) {
    expr_groups <- purrr::map(expr_groups, ~ formatR::tidy_source(text = .,
        comment = !remove_comments, arrow = TRUE, indent = 2, output = FALSE,
        width.cutoff = 50)) %>%
      purrr::map(getElement, "text.tidy") %>%
      purrr::map(~ readLines(textConnection(.)))
    for (i in seq_along(expr_groups)) {
      if (filesstrings::AllEqual(expr_groups[[i]], character(0))) {
        expr_groups[[i]] <- ""
      }
    }
  }
  empties <- purrr::map_lgl(expr_groups, ~ isTRUE(all.equal(., "")))
  expr_groups <- expr_groups[!empties]
  lapply(expr_groups, stringr::str_trim, side = "right")
  # str_trim because sometimes formatR leaves unnecessary trailing whitespace
}

#' Construct the shell of an `expect_equal` expression.
#'
#' Construct the shell an `expect_equal` expression from a character vector
#' containing an expression to be evaluated.
#'
#' @param text_expr A character vector of lines that, when executed produce an
#'   output.
#'
#' @return A character vector. The lines of text containing the `expect_equal`
#'   code (corresponding to the input `text_expr`), which will help to write the
#'   test file based on documentation examples. Remember that this is something
#'   that you're intended to fill the gaps in later.
#'
#' @examples
#' text_expr <- c("sum(1, ", "2)")
#' cat(paste(text_expr, collapse = "\n"))
#' construct_expect_equal(text_expr)
#' cat(paste(construct_expect_equal(text_expr), collapse = "\n"))
#' @export
construct_expect_equal <- function(text_expr) {
  text_expr[1] <- paste0("expect_equal(", text_expr[1])
  l <- length(text_expr)
  text_expr[l] <- paste0(text_expr[l], ", )")
  text_expr
}

#' Extract examples from a `.Rd` file as a character vector.
#'
#' This is a convenient wrapper to [tools::Rd2ex] which actually returns a character vector of the examples in the `.Rd` file.
#'
#' @param rd_file_path The path to the `.Rd` file.
#'
#' @return A character vector.
#'
#' @examples
#' this_function_rd <- system.file("extdata", "extract_examples_rd.Rd",
#'                                 package = "exampletestr")
#' extract_examples_rd(this_function_rd)
#' @export
extract_examples_rd <- function(rd_file_path) {
  tc <- textConnection(" ", "w")
  tools::Rd2ex(rd_file_path, tc)
  examples_lines <- textConnectionValue(tc)
  close(tc)
  examples_lines
}

```
So let's demonstrate `extract_examples`:
```{r Demonstrate extract_examples}
extract_examples("utils", pkg_dir = "tempkg")
```

Indeed we get all of the lines of the documentation examples. Now with `make_test_shell`, we turn it into something usable with `testthat`:
```{r Demonstrate make_test_shell}
lapply(extract_examples("utils", pkg_dir = "tempkg"), make_test_shell, "whatevs")
```
This might look a little weird in the output but it's really just 
```{r make make_test_shell look nice, eval=FALSE}
test_that("whatevs", {
  expect_equal(text_eval_error("a <- 1"), )
  expect_equal(text_eval_error("a <- "), )
})
```
and so on, which is what we would want. Now we have something we can fill in ourselves to create a real unit test.

We can make the unit tests _shell_ file (_shell_ because it needs to be filled in) via `make_tests_shell_file`. Running 
```{r make_tests_shells_file}
make_tests_shells_file("utils", pkg_dir = "tempkg")
```
outputs a `test_utils.R` file in the `tests/testthat` folder with contents
```{r test_utils.R contents, eval=FALSE}
test_that("construct_expect_equal works", {
  text_expr <- c("sum(1, ", "2)")
  expect_equal(cat(paste(text_expr, collapse = "\n")), )
  expect_equal(construct_expect_equal(text_expr), )
  expect_equal(cat(paste(construct_expect_equal(text_expr), collapse = "\n")), )
})

test_that("extract_examples_rd works", {
  this_function_rd <- system.file("extdata", "extract_examples_rd.Rd",
    package = "exampletestr")
  expect_equal(extract_examples_rd(this_function_rd), )
})

test_that("extract_expressions works", {
  text_expr <- c("a <- 1", "fx <- function(x) {", "  x + 1",
    "}  # this comment should disappear")
  expect_equal(extract_expressions(text_expr), )
})

test_that("text_parse_error works", {
  expect_equal(text_parse_error("a <- 1"), )
  expect_equal(text_parse_error("a <- "), )
})
```
which, for my purposes, I complete as
```{r fill in test shell}
test_that("text_parse_error works", {
  expect_false(text_parse_error("a <- 1"))
  expect_true(text_parse_error("a <- "))
})

test_that("extract_expressions works", {
  text_expr <- c("a <- 1",
                 "fx <- function(x) {",
                 "  x + 1",
                 "}")
  expect_equal(extract_expressions(text_expr), list(
    "a <- 1",
    c("fx <- function(x) {",
      "  x + 1",
      "}")
  ))
})

test_that("construct_expect_equal works", {
  text_expr <- c("sum(1, ", "2)")
  expect_equal(construct_expect_equal(text_expr), c("expect_equal(sum(1, ",
                                                    "2), )"))
})
```

To create these test shell files for each file in the `R/` directory of your package, run `make_tests_shells_pkg()`


```{r setdown, include=FALSE, echo=FALSE}
filesstrings::RemoveDirs("tempkg")
```

### The Goal is NOT Fully Automated Unit Test Creation
I would like to stress that whilst unit testing should be automatic, the creation of these tests is a manual process, a manual check. This package is supposed to help you *start* making those tests. It is not supposed to create fully functioning tests automatically, nor can it help you to write every type of test you might want.
