---
title: "exampletestr"
author: "Rory Nolan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{exampletestr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
knitr::opts_knit$set(root.dir = tempdir())
```

## In a nutshell
The primary use of `exampletestr` is from every file in the `R/` directory of a package, to create a corresponding _shell_ of a test file (something to get you started making tests (for use with `testthat`)) that will have to be filled in to create a fully functioning test file. This can be done for the whole package at once with `make_tests_shells_pkg()` or for individual files with `make_tests_shells_file()`. These tests are created by using the examples in the documentation of your package as a basis for tests. I have often (but not always) found that I test my functions manually by making sure my examples work as intended, so why not use these as a basis for proper unit testing?

## Out of the nutshell
First, let's set up a dummy package directory with just the `match.R` and `detect.R` files from the source code of the `stringr` package.
```{r setup}
library(exampletestr)
library(testthat)
devtools::create("tempkg")
devtools::use_testthat("tempkg")
file.copy(system.file("extdata", c("match.R", "detect.R"), package = "exampletestr"), 
          "tempkg/R")
devtools::document("tempkg")
```

The `match.R` file looks like this:
```{r Look at match.R file, eval=FALSE}
#' Extract matched groups from a string.
#'
#' Vectorised over `string` and `pattern`.
#'
#' @inheritParams str_detect
#' @param pattern Pattern to look for, as defined by an ICU regular
#'   expression. See [stringi::stringi-search-regex] for more details.
#' @return For `str_match`, a character matrix. First column is the
#'   complete match, followed by one column for each capture group.
#'   For `str_match_all`, a list of character matrices.
#'
#' @seealso [str_extract()] to extract the complete match,
#'   [stringi::stri_match()] for the underlying
#'   implementation.
#' @export
#' @examples
#' strings <- c(" 219 733 8965", "329-293-8753 ", "banana", "595 794 7569",
#'   "387 287 6718", "apple", "233.398.9187  ", "482 952 3315",
#'   "239 923 8115 and 842 566 4692", "Work: 579-499-7527", "$1000",
#'   "Home: 543.355.3679")
#' phone <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})"
#'
#' str_extract(strings, phone)
#' str_match(strings, phone)
#'
#' # Extract/match all
#' str_extract_all(strings, phone)
#' str_match_all(strings, phone)
#'
#' x <- c("<a> <b>", "<a> <>", "<a>", "", NA)
#' str_match(x, "<(.*?)> <(.*?)>")
#' str_match_all(x, "<(.*?)>")
#'
#' str_extract(x, "<.*?>")
#' str_extract_all(x, "<.*?>")
str_match <- function(string, pattern) {
  if (type(pattern) != "regex") {
    stop("Can only match regular expressions", call. = FALSE)
  }

  stri_match_first_regex(string,
    pattern,
    opts_regex = opts(pattern)
  )
}

#' @rdname str_match
#' @export
str_match_all <- function(string, pattern) {
  if (type(pattern) != "regex") {
    stop("Can only match regular expressions", call. = FALSE)
  }

  stri_match_all_regex(string,
    pattern,
    omit_no_match = TRUE,
    opts_regex = opts(pattern)
  )
}
```
So let's demonstrate `extract_examples`:
```{r Demonstrate extract_examples}
extract_examples("match", pkg_dir = "tempkg")
```

Indeed we get all of the lines of the documentation examples. Now with `make_test_shell`, we turn it into something usable with `testthat`:
```{r Demonstrate make_test_shell}
lapply(extract_examples("match", pkg_dir = "tempkg"), make_test_shell, "whatevs")
```

We can make the unit tests _shell_ file (_shell_ because it needs to be filled in) via `make_tests_shell_file`. Running 
```{r make_tests_shells_file}
make_tests_shells_file("match", pkg_dir = "tempkg")
```
```{r load stringr, include=FALSE}
library(stringr)
```
outputs a `test-utils.R` file in the `tests/testthat` folder with contents
```{r test-utils.R contents, eval=FALSE}
context("Match")

test_that("str_match() works", {
  strings <- c(
    " 219 733 8965", "329-293-8753 ", "banana", "595 794 7569",
    "387 287 6718", "apple", "233.398.9187  ", "482 952 3315",
    "239 923 8115 and 842 566 4692", "Work: 579-499-7527", "$1000",
    "Home: 543.355.3679"
  )
  phone <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})"
  expect_equal(str_extract(strings, phone), )
  expect_equal(str_match(strings, phone), )
  expect_equal(str_extract_all(strings, phone), )
  expect_equal(str_match_all(strings, phone), )
  x <- c("<a> <b>", "<a> <>", "<a>", "", NA)
  expect_equal(str_match(x, "<(.*?)> <(.*?)>"), )
  expect_equal(str_match_all(x, "<(.*?)>"), )
  expect_equal(str_extract(x, "<.*?>"), )
  expect_equal(str_extract_all(x, "<.*?>"), )
})
```
which can be sensibly completed as
```{r fill in test shell}
context("Match")

test_that("str_match() works", {
  strings <- c(
    " 219 733 8965", "329-293-8753 ", "banana", "595 794 7569",
    "387 287 6718", "apple", "233.398.9187  ", "482 952 3315",
    "239 923 8115 and 842 566 4692", "Work: 579-499-7527", "$1000",
    "Home: 543.355.3679"
  )
  phone <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})"
  expect_equal(str_extract(strings, phone), 
               c("219 733 8965", "329-293-8753", NA, "595 794 7569", 
                 "387 287 6718", NA, "233.398.9187", "482 952 3315", 
                 "239 923 8115", "579-499-7527", NA, "543.355.3679"))
  expect_equal(str_match(strings, phone), 
               matrix(c("219 733 8965", "219", "733", "8965",
                        "329-293-8753", "329", "293", "8753",
                        NA,             NA,    NA,    NA,    
                        "595 794 7569", "595", "794", "7569",
                        "387 287 6718", "387", "287", "6718",
                        NA,             NA,    NA,    NA,    
                        "233.398.9187", "233", "398", "9187",
                        "482 952 3315", "482", "952", "3315",
                        "239 923 8115", "239", "923", "8115",
                        "579-499-7527", "579", "499", "7527",
                        NA,             NA,    NA,    NA,    
                        "543.355.3679", "543", "355", "3679"), 
                      ncol = 4, byrow = TRUE))
  expect_equal(str_extract_all(strings, phone), 
               list("219 733 8965", "329-293-8753", character(0), 
                    "595 794 7569", "387 287 6718", character(0),
                    "233.398.9187", "482 952 3315", 
                    c("239 923 8115", "842 566 4692"), "579-499-7527",
                    character(0), "543.355.3679"))
  expect_equal(str_match_all(strings, phone), 
               list(t(c("219 733 8965", "219", "733", "8965")),
                    t(c("329-293-8753", "329", "293", "8753")),
                    matrix(character(0), ncol = 4),
                    t(c("595 794 7569", "595", "794", "7569")),
                    t(c("387 287 6718", "387", "287", "6718")),
                    matrix(character(0), ncol = 4),
                    t(c("233.398.9187", "233", "398", "9187")),
                    t(c("482 952 3315", "482", "952", "3315")),
                    matrix(c("239 923 8115", "239", "923", "8115",
                             "842 566 4692", "842", "566", "4692"),
                           ncol = 4, byrow = TRUE),
                    t(c("579-499-7527", "579", "499", "7527")),
                    matrix(character(0), ncol = 4),
                    t(c("543.355.3679", "543", "355", "3679"))))
  x <- c("<a> <b>", "<a> <>", "<a>", "", NA)
  expect_equal(str_match(x, "<(.*?)> <(.*?)>"),
               matrix(c("<a> <b>", "a",  "b", 
                        "<a> <>",  "a",  "",
                        NA,        NA,   NA,  
                        NA,        NA,   NA,  
                        NA,        NA,   NA), ncol = 3, byrow = TRUE))
  expect_equal(str_match_all(x, "<(.*?)>"), 
               list(matrix(c("<a>", "a", 
                             "<b>", "b"), ncol = 2, byrow = TRUE),
                    matrix(c("<a>", "a", 
                             "<>", ""), ncol = 2, byrow = TRUE),
                    t(c("<a>", "a")),
                    matrix(character(0), ncol = 2),
                    t(rep(NA_character_, 2))))
  expect_equal(str_extract(x, "<.*?>"), c("<a>", "<a>", "<a>", NA, NA))
  expect_equal(str_extract_all(x, "<.*?>"), 
               list(c("<a>", "<b>"), c("<a>", "<>"), "<a>", character(0),
                    NA_character_))
})
```

To create these test shell files for each file in the `R/` directory of your package, run `make_tests_shells_pkg()`.


```{r setdown, include=FALSE}
filesstrings::dir.remove("tempkg")
```

### The Goal is NOT Fully Automated Unit Test Creation
I would like to stress that whilst unit testing should be automatic, the creation of these tests is a manual process, a manual check. This package is supposed to help you *start* making those tests. It is not supposed to create fully functioning tests automatically, nor can it help you to write every type of test you might want.
