---
title: "One .R file at a time"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{One .R file at a time}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-setup, include=FALSE}
init_wd <- getwd()
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
knitr::opts_knit$set(root.dir = tempdir(check = TRUE))
pacman::p_load(testthat, exampletestr, stringr)
```

This vignette will teach you how to create a unit test shell for all of the examples in a single `.R` file's documentation.

I'm going to be extra careful and use `rprojroot` for my paths because I'm in the tricky situation of making a package inside a package vignette. If you're a package author and you don't know the `rprojroot` package, it's worth getting to know it.


## Setup

First, let's set up a dummy package called `tempkg` with just the `detect.R` and `match.R` files from the source code of the `stringr` package.

```{r setup, results='hide'}
usethis::create_package("tempkg", open = FALSE)
file.copy(
  system.file("extdata", c("detect.R", "match.R"), package = "exampletestr"), 
  rprojroot::find_root_file(
    "tempkg/R/", 
    criterion = rprojroot::has_dir("tempkg")
  )
)
```

Now let's move into the `tempkg` directory.

```{r fake-dir-change, eval=FALSE}
setwd(
  rprojroot::find_root_file(
    "tempkg/", 
    criterion = rprojroot::has_dir("tempkg")
  )
)
```

```{r further-knitr-setup, include=FALSE}
knitr::opts_knit$set(root.dir = paste0(tempdir(), "/", "tempkg"))
```

The `detect.R` file looks like this (well, I've abbreviated it slightly):
```{r Look at match.R file, eval=FALSE}
#' Detect the presence or absence of a pattern in a string.
#'
#' Vectorised over `string` and `pattern`.
#'
#' @param string Input vector. Either a character vector, or something
#'  coercible to one.
#' @param pattern Pattern to look for.
#'
#' @return A logical vector.
#' 
#' @export
#' @examples
#' fruit <- c("apple", "banana", "pear", "pinapple")
#' str_detect(fruit, "a")
#' str_detect(fruit, "^a")
#' str_detect(fruit, "a$")
#' str_detect(fruit, "b")
#' str_detect(fruit, "[aeiou]")
#'
str_detect <- function(string, pattern) {
  switch(type(pattern),
         empty = ,
         bound = str_count(string, pattern) > 0,
         fixed = stri_detect_fixed(string, pattern, opts_fixed = opts(pattern)),
         coll  = stri_detect_coll(string,  pattern, 
                                  opts_collator = opts(pattern)),
         regex = stri_detect_regex(string, pattern, opts_regex = opts(pattern))
  )
}
```

We can make the unit tests _shell_ file for the `str_match()` function using  `make_tests_shells_file()`:

```{r make_tests_shells_file, eval=FALSE}
make_tests_shells_file("detect", open = FALSE)
```

This outputs a `test-detect-examples.R` file in the `tests/testthat` folder with contents
```{r test-utils.R contents, eval=FALSE}
context("Detect")

test_that("`str_detect()` works", {
  fruit <- c("apple", "banana", "pear", "pinapple")
  expect_equal(str_detect(fruit, "a"), )
  expect_equal(str_detect(fruit, "^a"), )
  expect_equal(str_detect(fruit, "a$"), )
  expect_equal(str_detect(fruit, "b"), )
  expect_equal(str_detect(fruit, "[aeiou]"), )
  expect_equal(str_detect("aecfg", letters), )
})
```

which can be sensibly completed as

```{r fill in test shell}
context("Detect")

test_that("`str_detect()` works", {
  fruit <- c("apple", "banana", "pear", "pinapple")
  expect_equal(str_detect(fruit, "a"), rep(TRUE, 4))
  expect_equal(str_detect(fruit, "^a"), c(TRUE, rep(FALSE, 3)))
  expect_equal(str_detect(fruit, "a$"), c(FALSE, TRUE, FALSE, FALSE))
  expect_equal(str_detect(fruit, "b"), c(FALSE, TRUE, FALSE, FALSE))
  expect_equal(str_detect(fruit, "[aeiou]"), rep(TRUE, 4))
  expect_equal(str_detect("aecfg", letters), 
               letters %in% c("a", "c", "e", "f", "g"))
})
```


```{r setdown, include=FALSE}
knitr::opts_knit$set(root.dir = tempdir(check = TRUE))
if (dir.exists("tempkg")) filesstrings::dir.remove("tempkg")
knitr::opts_knit$set(root.dir = init_wd)
```


